---
title: "HW3-Fogelman"
output: html_document
---

##Kaelyn Fogelman 
###Homework 3 
###Due Date: 16 Oct 2020
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load packages necessary for the assignment.

```{r packages, include = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate)
```

##Covid Data 

Load and check dataset

```{r load data, include=TRUE}
covid <- read.csv("covid.csv", row.names = 1)
# Check data
str(covid)
```
<br><br>
1. Correctly format the dates in “Date.of.birth”, “First.day.of.symptoms”, “Date.of.outcome” and “Date.of.diagnosis”.
```{r format dates, include=TRUE}
covid$Date.of.birth <- 
  as.Date(parse_date_time(covid$Date.of.birth, "mdy"))

covid$First.day.of.symptoms <- 
  as.Date(parse_date_time(covid$First.day.of.symptoms, "mdy"))

covid$Date.of.outcome <- 
  as.Date(parse_date_time(covid$Date.of.outcome, "mdy"))

covid$Date.of.diagnosis <- 
  as.Date(parse_date_time(covid$Date.of.diagnosis, "mdy"))
```
<br><br>
2. Format “Hospitalization.type”, “Symptoms” and “Outcome” as factors and “Epidemiological.link…Notes” as character.

```{r format factor/character, include=TRUE}
covid$Hospitalization.type <- as.factor(covid$Hospitalization.type)

covid$Symptoms <- as.factor(covid$Symptoms)

covid$Outcome <- as.factor(covid$Outcome)

covid$Epidemiological.link...Notes <- as.character(covid$Epidemiological.link...Notes)
```
<br><br>
3. Create a three-dimensional table reporting the three factors from the previous question. Hint: use the function table() with the variables of interest.

```{r table, include = TRUE}
table1 <- table(covid[c("Hospitalization.type", "Symptoms", "Outcome")])
table1
```
<br><br>
4. Using a table object created in the previous question, subset it in order to create the following table: 

```{r subset table, include = TRUE}
table2 <- table1[1:2, 2, c(2:3)]
table2

##Gabii's code 

```
<br><br>
5. Order the data based on the date of diagnosis (from first to most recent).

```{r diagnosis order, include = TRUE}
covid_diagnosis.date <- arrange(covid, Date.of.diagnosis)
head(covid_diagnosis.date)
```

6. Add a column that reports whether or not a case was asymptomatic AND in home isolation. Name the observation “Home_Asymptomatic” if the conditions apply and “Non_Home_Asymptomatic” if not and then produce a bar plot of this new variable.

```{r new data, include = TRUE}
covid <- covid %>%
  mutate(Asymptomatic_AND_Home_Isolation = 
           ifelse(covid$Symptoms == "Asymptomatic" & 
                  covid$Hospitalization.type == "Home isolation", 
                  "Home_Asymptomatic", "Non_Home_Asymptomatic"))
counts <- table(covid$Asymptomatic_AND_Home_Isolation)
barplot(counts, main="Asymptomatic_AND_Home_Isolation",
   xlab="Symptom status and Isolation status")
```
<br><br>
7. Count the number of cases of people born after 1981 and that have healed.
```{r 1981healed, include = TRUE}
covid %>%
  filter(Date.of.birth > ymd(19810101) & Outcome == "Healed") %>%
  count()
```
<center> __158 people that are were born after 1981 have healed.__ </center>
<br><br>

8. Count the number of cases that are asymptomatic OR in home isolation (but not both) AND were born before 1982.

```{r symptom&hospitalization.type, include = TRUE}
covid %>%
  filter(Symptoms == "Asymptomatic" | Hospitalization.type == "Home isolation"
         & Date.of.birth < ymd(19820101)) %>%
  count()
```
<center> __273 people born before 1982 are asymptomatic or in home isolation.__ </center> <br><br>

9. Create a new dataset including only the rows where “Epidemiological.link…Notes” includes the words “contact” OR “symptom” (or both). Hint: you can use the grep() function and tolower().

```{r dataset, inclue = TRUE, warning = FALSE}
covid_new_data <- covid %>% filter (grepl("contact", pattern = 
                              tolower(covid$Epidemiological.link...Notes)) | 
                              grepl("symptom", tolower(covid$Epidemiological.link...Notes)))
head(covid_new_data)
```
<br><br>
10. In the previous dataset add a column reporting the age (in years, therefore in integer format) of each patient as of October 2nd, 2020. Save this dataset into a .csv file and make it available on your GitHub repository for this assignment.

```{r new column, include = TRUE}
covid_new_data <- covid_new_data %>%
  mutate("Age" = as.period(interval(covid_new_data$Date.of.birth, 
      ymd("2020 October 2")))$year) %>%
           write_csv("hw3.kf.new_data_set.csv")
```
<br><br>
11. Produce a pie chart for the type of hospitalization for cases born between 1960 and 1980.

```{r piechart, include = TRUE}
library(ggplot2)
pie_covid_df <- subset(covid, covid$Date.of.birth >= as.Date("1960-01-01") & 
         covid$Date.of.birth <= as.Date("1980-01-01"))

ggplot(pie_covid_df, aes(x=factor(1), fill=Hospitalization.type))+
    geom_bar(width = 1)+
    coord_polar("y")
```
<br><br>

##Salary 

1. Load the data and check the formatting of the variables.

```{r load data2, include=TRUE}
adult <- read.csv("adult.csv", row.names = 1)
# Check data
str(adult)
```
<br><br>
2. Rename the last column (currently called “NA”) containing the dichotomous salary information. Assign the name salary to it.
```{r rename salary, include = TRUE}
adult <- rename(adult, salary = 15)
head(adult)
```
<br><br>
3. The values in salary have a space in front of them (e.g. " >50K"): remove the space from all values (hint: you can use substring()).
```{r remove space, inclue = TRUE}
adult$salary <- trimws(adult$salary, which = c("left"))
```
<br><br>
4. Again, in the variable salary, replace “>50K” with the value 1 and “<=50K” with 0. Make sure to format it as a factor with two levels.
```{r normalize salary, include = TRUE}
adult2 <- adult %>% 
  mutate(salary = recode(salary, ">50K" = "1",
                                   "<=50K" = "0"))
adult2$salary <- as.numeric(adult2$salary)
```
<br><br>
5. Use the glm() function to estimate this logistic model. Only specify the arguments formula = salary ~ ., data and family = binomial. Save the result of the estimation in an object called fit.

```{r glm, include = TRUE, warning = FALSE, message = FALSE}
fit = glm(formula = salary ~ ., data = adult2, family = binomial)
```
<br><br>
6. Using the information in fit (hint: you can use the functions coef() and summary() to extract information), create a dataframe collecting:<br>
  6.1 The names of the variables (names of the rows of the dataframe);<br>
  6.2 The value of the coefficients (first column);<br>
  6.3 A logical vector stating which coefficients are positive (second column);<br>
  6.4 The p-values (third column).

```{r fit dataframe, include = TRUE, warning = FALSE, error = FALSE}
##Variable names, column 1 
df <- data.frame (row.names1 = (as.data.frame(coef(fit))%>%
              drop_na() %>%
              row.names() ))
                          
##Coefficient value, column 2 
df$coef.value <- drop_na (as.data.frame(coef(fit)))[[1]]

##Positive coefficient = TRUE, negative coefficient = FALSE, column 3 
df <- df %>%
  mutate("positive.coef" = as.logical(coef.value > 0))

##p-value, column 4 
df$p.value <- as.data.frame(summary(fit)[["coefficients"]])$'Pr(>|z|)'

head(df)
``` 
<br><br>
7. Subset the dataframe created in the previous question to only show the rows where the p-values are strictly smaller than 0.05. Knowing that (i) the remaining rows are statistically significant and that (ii) positive coefficients contribute to increase the probability of a salary larger than $50,000 (the opposite for negative values), comment on those variables that negatively contribute to salary.

```{r }
sigvar <- df %>% 
  filter (df$p.value < 0.05)
sigvar
```

<center> __Variables that contribute negatively to salary include: education 7th-8th ,  marital.status Never-married, occupation Farming-fishing, occupation Handlers-cleaners, occupation Other-service, occupation Priv-house-serv,  relationship Own-child,  native.country Columbia, native.country South.__<br><br>

__Based on this analysis, education impacts salary with indivudals having a lower education level (7th-8th grade) having their salary negatively impacted compared to individuals that have higher education levels.__ <br><br>

__Individuals that have never been married have their salary negatively impacted compared to individuals who are currently married or have been married previously and are separated or widowed.__ <br><br>

__Individuals who have occupations that are labor, service or domestic based have their salaries more negatively impacted to individuals who have more technical or professional occupations.__ <br><br>

__Considering family relationships, having children in the household impacts salary more significantly than other family relationships including marriage or having a non-child relative in the household.__ <br><br>

__Those from the countries of Columbia and "South" (could indicate South Korea or South Africa - not indicated in the data set) have their salaries more negatively influenced than individuals with other countries of origin.__ <br><br>
</center>

## Bonus 

```{r packages bonus, include = TRUE, warning = FALSE, message = FALSE}
library(leaflet)
library(htmltools)
```
<br><br>
```{r data for map, include = TRUE}
teams <- c("Real Madrid", "FC Barcelona", "Atletico Madrid", 
           "Atletico Bilbao", "Valencia")
champions <- c(34, 26, 10, 8, 6)
city <- c("Madrid", "Barcelona", "Madrid", "Bilbao", "Valencia")

map.df <- cbind(teams, champions, city)

head(map.df)
```
<br><br>
Create base map 
```{r create base map, include = TRUE}
map <- leaflet() %>% 
  setView(lng = 0.8891, lat = 41.6488, zoom = 4) %>%
  addTiles()
map
```
<br><br>
```{r popups and circles, include = TRUE}
content <- paste(sep = "<br/>",
                 "<b>Real Madrid</b>",
                 "34", 
                 "<b>Atletico Madrid</b>",
                 "10")

content2 <- paste(sep = "<br/>",
                 "<b>FC Barcelona</b>",
                 "26")

content3 <- paste(sep = "<br/>",
                 "<b>Athletic Bilbao</b>",
                 "8")

content4 <- paste(sep = "<br/>",
                 "<b>Valencia</b>",
                 "6")

map <- map %>%
  leaflet() %>%
  addTiles() %>%
  addPopups (-3.7038, 40.4168, content, 
             options = popupOptions(closeButton = FALSE)) %>%
  addPopups (-2.1734, 41.3851, content2, 
             options = popupOptions(closeButton = FALSE)) %>%
  addPopups (-2.9350, 43.2630, content3, 
             options = popupOptions(closeButton = FALSE)) %>%
  addPopups (-0.3763, 39.4699, content4, 
             options = popupOptions(closeButton = FALSE)) %>%
  addCircles (lng = -3.7038, lat = 40.4168, weight = 5, 
              radius = (34000 * 2)) %>%
  addCircles (lng = -3.7038, lat = 40.4168, weight = 5, 
              radius = (10000 * 2)) %>%
  addCircles (lng = -2.1734, lat = 41.3851, weight = 5, 
              radius = (26000 * 2)) %>%
  addCircles (lng = -2.9350, lat = 43.2630, weight = 5, 
              radius = (8000 * 2)) %>%
  addCircles (lng = -0.3763, lat = 39.4699, weight = 5, 
              radius = (6000 * 2)) 
 
map
```
